%\documentclass[sigconf]{acmart}
%\acmConference[IFL'17]{International Symposium on Implementation and Application of Functional Languages}{August 2017}{Bristol, United Kingdom}
%\acmYear{2017}
%\copyrightyear{2017}

\documentclass[sigconf]{acmart}
\acmConference[IFL'17]{International Symposium on Implementation and Application of Functional Languages}{August 2017}{Bristol, United Kingdom}
\acmYear{2018}
\copyrightyear{2018}

%The ISBN number assigned to IFL 2017 is 978-1-4503-6343-3, and can be entered into the template as:

\acmISBN{978-1-4503-6343-3}


%\documentclass[sigplan,10pt]{acmart}

%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

\bibliographystyle{ACM-Reference-Format}
\usepackage{flushend} 
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage[export]{adjustbox}
\usepackage{graphicx}
\usepackage{fancyvrb}

\usepackage{amsmath}
\usepackage{latexsym}
%\usepackage[usenames,dvipsnames]{color}

\usepackage{url}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
%\input{shadded}

\newcounter{figpart}[figure]
\renewcommand{\thefigpart}{}%%\alph{figpart}}

\usepackage[utf8]{inputenc}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{bashful}
\usepackage{setspace}
\usepackage{color}
\usepackage{xcolor}
%% \lstloadlanguages{Erlang}
%% \lstset{
%%   frame=tb,
%%   rulecolor=\color{gray},
%%   basicstyle=\small\ttfamily\setstretch{0.8},
%%   breakindent=4ex,
%%   postbreak=(cont.),
%%   breaklines=true,
%%   showspaces=false,
%%   showtabs=false,
%%   xleftmargin=15pt,
%%   framexleftmargin=5pt,
%%   framexrightmargin=0pt,
%%   framexbottommargin=0pt,
%%   captionpos=b,
%%   numbers=left,
%%   numberstyle=\tiny\color{gray},
%% }
%% \hypersetup{
%%     colorlinks,
%%     linkcolor={red},
%%     citecolor={blue!50!black},
%%     urlcolor={blue!80!black}
%% }

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codeblue}{rgb}{0.0,0,0.62}
\definecolor{backcolour}{rgb}{0.95,0.95,0.95}

\newcommand{\lstbasicfont}{\small\fontfamily{pcr}\selectfont}
\newcommand{\lstcommentstyle}{\color{codegreen}\itshape}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\lstcommentstyle,
    keywordstyle=\color{codeblue}\bfseries,
    %identifierstyle=\color{black},
    identifierstyle=\color{codegray}\bfseries,
    %numberstyle=\tiny\color{codegray},
    stringstyle=\color{purple},
    basicstyle=\lstbasicfont,
    %basicstyle=\small\ttfamily,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    %numbers=left,
    %numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}

\lstdefinelanguage{myerlang}[]{erlang}{
  morekeywords={spec,fun,bitstring,orelse,record,include_lib},
  otherkeywords={|,||,<-,->,!,[,],\{,\}},
  deletekeywords={list,ok}
}

\lstset{style=mystyle,
  literate=
    {∪}{{$\cup$}}1 {∩}{{$\cap$}}1 {∈}{{$\in$}}1 {⊆}{{$\subseteq$}}1
    {∘}{{$\circ$}}1 {-}{{-}}1 {~}{{\texttildelow}}1 {<}{{<}}1
    {>}{{>}}1
  }

%\newcommand{\edge}[1]
%           {\stackrel{\bf{#1}}{\rightarrow}}
\newcommand{\edge}[1]
           {\xrightarrow{\bf{#1}}}
\newcommand{\rel}[1]
           {\stackrel{\bf{#1}}{\leadsto}}

\newcommand{\dc}{\textit{d\&c}}


\begin{document}

\title[Pattern Candidate Discovery and Parallelization Techniques]{Pattern Candidate Discovery and \\ Parallelization Techniques}

\author{Melinda T\'oth, Istv\'an Boz\'o, Tam\'as Kozsik}
%\authornote{with author1 note}
\orcid{nnnn-nnnn-nnnn-nnnn}
\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
  \institution{ELTE, E\"otv\"os Lor\'and University}            %% \institution is required
%  \streetaddress{Street1 Address1}
  \city{Budapest}
%  \state{State1}
%  \postcode{Post-Code1}
  \country{Hungary}                    %% \country is recommended
}
\email{{tothmelinda,bozoistvan,kto}@elte.hu}          %% \email is recommended

%% \author{Melinda T\'oth, Istv\'an Boz\'o, Tam\'as Kozsik}


\begin{abstract}
Parallel computations in a program can be expressed conveniently,
at a high level of abstraction, using parallel patterns such as
task farm, pipeline or divide-and-conquer. In order to transform a sequential
program into a pattern-based parallel one, the software developer
may want to apply refactoring transformations on it. This tutorial
explains a methodology to perform tool supported parallelization of
programs by presenting how to use a specific static source code
analysis and transformation system for Erlang. A key element of the
approach is pattern candidate discovery, a static analysis technique
to identify code fragments that can be refactored into a parallel
pattern.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011073</concept_id>
<concept_desc>Software and its engineering~Software maintenance tools</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011074.10011092</concept_id>
<concept_desc>Software and its engineering~Software development techniques</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software maintenance tools}
\ccsdesc[500]{Software and its engineering~Software development techniques}

%% End of generated code


%% Keywords
%% comma separated list
\keywords{pattern based programming, Erlang, PaRTE, static analysis, refactoring}  %% \keywords are mandatory in final camera-ready submission


%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
\input{intro}

\section{Example: a pattern candidate}
\input{example}

\input{patterns}
%\input{skel}


\section{Tool support for parallelisation}

\input{parte-tool}

\section{How to build a tool like PaRTE?}
\input{analysis}


\section{Evaluation}
\input{eval}

\section{Related work}
\input{related}


\section{Summary}

The ParaPhrase Refactoring Tool for Erlang facilitates the parallelization
of Erlang source code by providing automated pattern candidate discovery and
refactoring transformations. This paper presents the underlying methodology,
and teaches the use of the tool. Moreover, we explain how the static source
code analyses work for parallel pattern candidate discovery -- this information
can be used to apply the methodology, and implement the parallelization tool,
in the context of other programming languages. Currently we are working on
a refactoring tool-chain for Scala. This language has a concurrency tool-set (Akka)
very similar to that of Erlang. Therefore, the main ideas presented in this
paper can be reused in the parallelisation of Scala programs -- though many interesting
problems remain due to the differences in the two languages. Even more
problems can arise when targeting other programming languages (e.g. OpenMP
would be a logical choice).

This presentation focused on a set of well-known parallel patterns, including
the task farm, the pipeline and the divide-and-conquer patterns. We explained
what kind of program fragments (pattern candidates) can be automatically
identified by PaRTE, and how to select the most promising candidates.
The selected candidates can be refactored into
applications of an algorithmic skeleton library, which implements the
parallel patterns. Many of the refactoring transformations can be performed
automatically by the tool, though the most complex ones can only be carried
out by hand currently. We have evaluated the applicability of the methodology
on a number of real-world, open source code bases, and found that the tool
is indeed able to identify many interesting parallel pattern candidates.

There are still, of course, some shortcomings and unsolved implementation
issues. We do not claim that our parallel pattern candidate discovery is
either complete or even sound. Some of the discovery rules are trade-offs
between soundness and (efficient) computability. It might happen that the
tool presents a ``fake candidate'', which cannot be refactored into a pattern.
This should not be a problem, however, since our tool is not a compiler
optimization running unsupervised. Fake candidates can be discarded by the
user after proper consideration.

The completeness of the discovery is not guaranteed either. Our rules can
find many interesting candidates, but there may be many more: in the future
the rule set can be extended, or can be made more accurate. What is even
more important is the good ranking of the candidates. Completely automatic
performance estimations are not feasible in more complex cases, so either
good heuristics are required (such as our existing one using some complexity
metrics), or the integration  of our tool with an environment supporting
development-time experimentation and dynamic analyses.

Automatic transformations also have a number of unsolved issues. It might
happen that a candidate is found by the discovery, but the current
definition of the appropriate transformation is not able to cover that
specific candidate. In this case either manual transformation of the code
is needed (which is error-prone) or the application of a sequence of
built-in transformations including program shaping before, and code clean-up
after, the introduction of the desired pattern (like in the case of the \dc{}
pattern). We are working actively to ensure, and also to prove, soundness
of the refactoring transformations -- clearly, this is a crucial requirement
for our approach.


\subsection*{Acknowledgement}
The research has been supported by the European Union, co-financed by the European Social Fund (EFOP-3.6.2-16-2017-00013) in 2017-2018, 
by the Hungarian Government through the New National Excellence Program of the Ministry of Human Capacities in 2016-2017, 
by the Higher Education Restructuring Fund of the Hungarian Government in 2015-2016, 
and by the Seventh Framework Programme (FP7) of the European Union under the contract number 288570 in 2013-2015.

\bibliography{ifl17}

%% \appendix
%% \setcounter{lemma}{0}
%% \section{The detailed proofs of lemmas.}\label{proofs}
%% \input{lemmas.tex}

\end{document}
